---
title: "CodeBook"
output: html_document
---

Variable list and descriptions
------------------------------

Variable name    | Description
-----------------|------------
subject          | ID the subject who performed the activity for each window sample. Its range is from 1 to 30.
activity         | Activity name
featDomain       | Feature: Time domain signal or frequency domain signal (Time or Freq)
featInstrument   | Feature: Measuring instrument (Accelerometer or Gyroscope)
featAcceleration | Feature: Acceleration signal (Body or Gravity)
featVariable     | Feature: Variable (Mean or SD)
featJerk         | Feature: Jerk signal
featMagnitude    | Feature: Magnitude of the signals calculated using the Euclidean norm
featAxis         | Feature: 3-axial signals in the X, Y and Z directions (X, Y, or Z)
featCount        | Feature: Count of data points used to compute `average`
featAverage      | Feature: Average of each variable for each activity and each subject

DataSet Structure
```{r}
str(DataTidy)
```
List variables in Data  
```{r}
colnames(DataTidy)
```
Show a few rows of dataset
```{r}
DataTidy
```
Summary of Variables
```{r}
summary(DataTidy)
```

Save to File
```{r}
f <- file.path(path, "result.txt")
write.table(DataTidy,f, row.name=FALSE)
```